\herschel\footnote{{\it Herschel} is an ESA space observatory with science instruments provided by European-led Principal Investigator consortia and with important participation from NASA.} Orion Protostar Survey (HOPS) Imaging and Photometry Results.

### 2018-01-20 HOPS/photometry/ms.tex
all changes in

\begin{abstract}
We summarize our investigation of the observed far-infrared properties of 410 protostar candidates in the Orion~A and B clouds using data from the \herschel\ Space Observatory.  The observations were taken as part of the Open Time Key Program {\it HOPS} (\herschel\ Orion Protostar Survey, PI: S. T. Megeath).  The photometry portion of HOPS, discussed here, focused on 70 and 160~\micron photometry with the \herschel/PACS instrument.  This work presents our observations, data processing and map-making, and photometry analyses using both aperture and point spread function fitted photometry.
\par
We find spatial correlations in the observed flux ratios as a function of sub-regions in both the Orion~A and B clouds.  Using simulations, we rule out completeness bias as the sole cause for the observed correlations.  We offer differences in accretion properties and/or evolutionary status as, at least partially, responsible for the observed differences by sub-regions.  Both accretion properties and evolutionary status differences, in turn, may trace fundamental differences in the star-forming conditions in the respective environments.  \end{abstract}

@@

Unfortunately, this strong link to their surroundings also makes protostars difficult to study.  In particular:  (i) observational studies must be able to disentangle the degenerate effects of both the environment and the star-formation process itself.  Protostars are characterized by a central core, an accretion disk, outflows, and an infalling envelope \citep{ali}. Additionally, protostars alter their local environment with phenomenon such as Herbig-Haro objects \citep{hhobjects}.  The spatial scales of these components range from sub-AU to 100s of AUs \citep{mo07}.  Thus, observational strategies must account for this compact, yet complicated morphology.  (ii) The presence of dust in the local environment means the amount of extinction in lines-of-sight to star formation regions is abnormally high (few to several tens of magnitudes in the visual) compared to most of the rest of the Galaxy \citep{dobbs}.  Finally, (iii)  \spitzer\ and \herschel\ observations, in particular, confirmed long recognized views that far-IR observations of the reprocessed radiation can probe the evolution of protostars \citep{als87}.  The presence of an infalling envelope of gas and dust is a defining characteristic of protostars. This envelope absorbs the radiation from the central accreting star+disk and reprocesses most of the luminosity into the far-IR \citep{ali}.   Far-IR observations are, therefore, absolutely crucial for efficiently identifying and studying protostars because protostars emit most of their light at those wavelengths.

@@

\section{Protostar sample \& observations}
\label{sec:sample}
\par
The \spitzer\ Orion protostar sample was defined and described by \cite{orion}.  For HOPS, we selected a flux-limited subset of 372 protostars for follow-up with the \herschel\ Photodetector Array Camera and Spectrometer (PACS) instrument \cite{pacs}.  The  HOPS targets are spatially spread over the Orion A and B molecular clouds from declinations of $-8^\circ50'$ to $1^\circ54'$ and from right ascensions of $5^{\rm h}33^{\rm m}$ to $5^{\rm h}55^{\rm m}$.  The \herschel\ observations were designed to detect those protostars with an expected 70~\micron\ flux in excess of 42 mJy.  The number of targets for the HOPS follow-up was limited solely by signal-to-noise ratio (S/N) considerations.   Thirty three (33) objects in this initial list were not observed with \herschel. An additional four are duplicates of existing sources, in which scattered light detected with Spitzer/IRAC turned out to be associated with a nearby point source.  See \citet{furlan} for further details.
\par
Our original sample of 372 \herschel\ targets were segregated in distinct spatial tiles to optimize observing.  The tiles were assigned consecutive three-digit integer group numbers starting with the number 0 (zero).  The spatial sizes of the tiles ranged from 3\arcmin\ (most common) to 7\arcmin).  The protostar candidates selected for our \herschel\ survey are referred to as the HOPS sources and are uniquely identified by a 3-digit integer number starting with the number 0 (zero).   Ultimately, we detected additional 22 candidates previously identified in the \spitzer\ survey whose spatial location happened to be within the field-of-view of the observation tile.  These sources, however, were not part of the original HOPS candidate list because their estimated fluxes were below our S/N limit.  Further, \cite{pbrpaper} found protostars and candidates not identified in the \spitzer\ list; See section~\ref{sec:protostars} for more details.  Our final catalog appends the original sample of 372 targets with the additions mentioned above.  There are a total of 410 sources in the final sample included here.
\par
\cite{furlan}\ classified 330 of these sources as young stellar objects (YSOs) and the remaining as unlikely to be protostars and/or sources not observed or detected by PACS at 70~micron.  For their analysis, \cite{furlan} focused on the 330 HOPS targets which have complete \spitzer\ and \herschel\ data (at least a PACS 70~\micron\ detection) and most were originally considered protostars by \cite{orion}.  The objective of this contribution is to provide complete data on all sources from our original sample of 372 protostars as well as the additional targets detected in our images.  Thus, we have included all 410 objects in this contribution.

@@

\section{Results}
\label{sec-results}
\par
In this contribution, we focus on the {\em observed}\ photometric properties of the target protostars, focusing on the fluxes in four bands: the 24~\micron\ from \spitzer/MIPS and the 0, 100, and 160~\micron\ bands from \herschel/PACS instruments..  A series of companion publications, \cite{will}, \cite{furlan}, \cite{erin}, and \cite{pbrpaper} present results and analysis from other scientific investigations and interpretations using these data from the HOPS program.   Our analysis concentrates on the spatial variations in the flux and color distributions across the Orion clouds.  To search for variations in the color and flux distributions, we divide the Orion clouds into X distinct sub-regions, shown in Table~\ref{tbl:regions}.  We augment our analysis with previously published results for the PBR sources \cite{pbrpaper}, and sub-mm photometry \citep{thomas}.  We present results from each of the above identified observation quantity below, and collectively discuss the implications for star-formation implied by these observations in Section~\ref{sec:discussion}.

@@

\subsection{Flux ratio distribution}
\label{sec:fratio}
\par
Flux ratios (colors) are better suited for determining differences as a function of source type than the flux itself because flux ratios depend on the shape of the source SED \citep{ali}. We investigated two different flux ratios using the combined Spitzer, HOPS and Gould-Belt photometry: \clra, and \clrb. Figure~\ref{fig:clrclr} shows the flux-ratio vs flux-ratio plots segregated by sub-region for Orion A and Orion B. We do not include data with upper limits for these plots (sources with flag value = 2 in Table~\ref{tbl:obs}). The red symbols show the colors for the sub-region identified in the sub-panel, and the gay symbols show the colors for the entire sample. As with the median of the FD, we note an apparent systematic change in the observed flux ratios relative to the entire population in both Orion~A and Orion~B.
\par
We quantitatively explore this apparent trend further in Figures~\ref{fig:clr1_v_dec}~\&~\ref{fig:clr2_v_dec}. These Figures show the mean and median colors for each of the sub-regions as a function of the mean declination for all sources belonging to the respective sub-region.
\par
In Orion~A, the median and mean colors of the sub-regions become systematically redder towards higher declinations.  In Orion~B, the median and mean colors of the sub-regions becomes systematically redder towards lower declinations.  For both, the vertical bar shows the 1-$\sigma$\ standard deviation of the flux-ratio values as a proxy for the range.  Even with large variation in color, the systematic trends with declination is clearly delineated.

@@

\section{Discussion}
\label{sec:discussion}
\par
The results presented in Section~\ref{sec-results} show the global properties of the HOPS protostar sample.  Since the sub-regions in Orion~A and B are well stratified by declination, differences are easily explored by using declination as a proxy for sub-region.   Indeed, Figures~\ref{fig:mdFlux_v_dec},~\ref{fig:clr1_v_dec},~\&~\ref{fig:clr2_v_dec} show that systematic variations exist in the observed proto-stellar properties as a function of declination in the Orion~A~\&~B clouds.  We discuss in Sections~\ref{sec:compbias}~\&~\ref{sec:popdiff} two explanations of this observed trend.   Again, we will restrict ourselves to observed quantities only and leave the detailed exploration of inferred properties to companion papers.  We rule out inclination angle (to the observer) as a factor because no reasonable justification can be made for systematic differences in inclination as a function of sub-region in Orion.

@@

We can now further investigate completeness limit and observational bias via a simple simulation.  The LDN 1641 sub-regions contain the largest number of protostars (see Table~\ref{tbl:regions}) and visual inspection shows the least contamination of the sample from extended nebulosity. Further, fainter protostars are detected in LDN 1641 than in Orion OMC 2/3. Hence, if bright, complex, spatially extended emission from the local dust is responsible for decreased reliability and higher completeness limits for point source photometry, then LDN 1641 provides an ideal contrast to the Orion OMC 2/3 sub-region. The contrast between LDN~1641 and Orion OMC-2/3 is at the extreme ends of the spatial correlation observed in Figures~\ref{fig:mdFlux_v_dec}$-$~\ref{fig:clrclr}.

@@

\subsection{Population differences}
\label{sec:popdiff}
\par
We also consider systematic differences in the underlying protostar population as a possible cause for the observed correlations in Figures~\ref{fig:clr1_v_dec}~\&~\ref{fig:clr2_v_dec}.  \cite{will} provide observational evidence for such a scenario.  \cite{will} derived bolometric luminosities and temperatures of our HOPS protostar sample and find statistical differences in the bolometric properties of the protostars as a function of the sub-region in Orion~A~and~B clouds.  The authors argue that these differences track an underlying difference in the star-formation rate across the sub-regions leading to a different mixture of protostars and more evolved objects.  Their findings are supported, in part, by evidence that the fractions for both Class~0 stars and PBRs compared to Class~I~and~II objects are also different in the subregions \citep{pbrpaper}.
\par
Indeed, \cite{ali} show that \herschel\ colors are highly correlated with the assumed envelope properties, especially, the infall rate and the envelope mass. In fact, \cite{ali} find protostellar envelopes to be the dominant factor affecting the observed colors in the \herschel\ bands.  Given that large, active envelopes are the defining characteristics of Class~0 objects and PBRs, the correlations apparent in Figures~\ref{fig:clr1_v_dec}~\&~\ref{fig:clr2_v_dec} are consistent with the population differences suggested by \cite{will}.

\section{Conclusions}
\label{sec:conc}
We explored the mid- to far-infrared properties of the Orion protostars obtained as part of the \herschel Key Program HOPS. Investigations of the observed properties of these protostars reveal apparent correlations as a function of sub-regions in both the Orion A and B clouds.  We investigated the role of bias introduced due to differences in completeness limits in these correlations by artificially introducing a completeness limit in the LDN~1641 sub-region and comparing the resulting changes in median flux ratio to Orion~OMC-2/3 values. These simulations suggest that completeness limit bias alone is not sufficient to explain the observed correlation for \clrb.  An alternative explanation may be population differences between these sub-regions. We offer
differences in accretion properties and/or evolutionary status as the cause for the population differences. We note that both accretion properties and evolutionary status trace fundamental differences in the star-forming conditions in the environment, notably higher gas+dust content.  This contribution focused entirely on the observed photometry and flux ratios.   Companion publications will explore the environmental differences by examining derived quantities such as bolometric properties and classification.

>>>>

\begin{abstract}
We summarize our investigation of the observed far-infrared properties of 410 protostar candidates in the Orion~A and B clouds using data from the \herschel\ Space Observatory.  The observations were taken as part of the Open Time Key Program {\it HOPS} (\herschel\ Orion Protostar Survey, PI: S. T. Megeath).  The photometry portion of HOPS, discussed here, focused on 70 and 160~\micron\  photometry with the \herschel/PACS instrument.  This work presents our observations, data processing and map-making, and photometry analyses using both aperture and point spread function fitted photometry.
\par
We find spatial correlations in the observed flux ratios as a function of sub-regions in both the Orion~A and B clouds.  Using simulations, we rule out completeness bias as the sole cause for the observed correlations.  We offer differences in accretion properties and/or evolutionary status as, at least partially, responsible for the observed differences by sub-regions.  Both accretion properties and evolutionary status differences, in turn, may trace fundamental differences in the star-forming conditions in the respective environments.  \end{abstract}

@@

Unfortunately, this strong link to their surroundings also makes protostars difficult to study.  In particular:  (i) observational studies must be able to disentangle the degenerate effects of both the environment and the star-formation process itself.  Protostars are characterized by a central core, an accretion disk, outflows, and an infalling envelope \citep{ali}. Additionally, protostars alter their local environment with phenomena such as Herbig-Haro objects \citep{hhobjects}.  The spatial scales of these components range from sub-AU to 100s of AUs \citep{mo07}.  Thus, observational strategies must account for this compact, yet complicated morphology.  (ii) The presence of dust in the local environment means the amount of extinction in lines-of-sight to star formation regions is abnormally high (few to several tens of magnitudes in the visual) compared to most of the rest of the Galaxy \citep{dobbs}.  Finally, (iii)  \spitzer\ and \herschel\ observations, in particular, confirmed long recognized views that far-IR observations of the reprocessed radiation can probe the evolution of protostars \citep{als87}.  The presence of an infalling envelope of gas and dust is a defining characteristic of protostars. This envelope absorbs the radiation from the central accreting star+disk and reprocesses most of the luminosity into the far-IR \citep{ali}.   Far-IR observations are, therefore, absolutely crucial for efficiently identifying and studying protostars because protostars emit most of their light at those wavelengths.

@@

\section{Protostar sample \& observations}
\label{sec:sample}
\par
The \spitzer\ Orion protostar sample was defined and described by \cite{orion}.  For HOPS, we selected a flux-limited subset of 372 protostars for follow-up with the \herschel\ Photodetector Array Camera and Spectrometer (PACS) instrument \cite{pacs}.  The  HOPS targets are spatially spread over the Orion A and B molecular clouds from declinations of $-8^\circ50'$ to $1^\circ54'$ and from right ascensions of $5^{\rm h}33^{\rm m}$ to $5^{\rm h}55^{\rm m}$.  The \herschel\ observations were designed to detect those protostars with an expected 70~\micron\ flux in excess of 42 mJy.  The number of targets for the HOPS follow-up was limited solely by signal-to-noise ratio (S/N) considerations.   Thirty three (33) objects in this initial list were not observed with \herschel. An additional four are duplicates of existing sources, in which scattered light detected with Spitzer/IRAC turned out to be associated with a nearby point source.  See \citet{furlan} for further details.
\par
Our original sample of 372 \herschel\ targets was segregated in distinct spatial tiles to optimize observing.  The tiles were assigned consecutive three-digit integer group numbers starting with the number 0 (zero).  The spatial sizes of the tiles ranged from 3\arcmin\ (most common) to 7\arcmin).  The protostar candidates selected for our \herschel\ survey are referred to as the HOPS sources and are uniquely identified by a 3-digit integer number starting with the number 0 (zero).   Ultimately, we detected additional 22 candidates previously identified in the \spitzer\ survey whose spatial location happened to be within the field-of-view of the observation tile.  These sources, however, were not part of the original HOPS candidate list because their estimated fluxes were below our S/N limit.  Further, \cite{pbrpaper} found protostars and candidates not identified in the \spitzer\ list; See section~\ref{sec:protostars} for more details.  Our final catalog appends the original sample of 372 targets with the additions mentioned above.  There are a total of 410 sources in the final sample included here.
\par
\cite{furlan}\ classified 330 of these sources as young stellar objects (YSOs) and the remaining as unlikely to be protostars and/or sources not observed or detected by PACS at 70~micron.  For their analysis, \cite{furlan} focused on the 330 HOPS targets which have complete \spitzer\ and \herschel\ data (at least a PACS 70~\micron\ detection) and most were originally considered protostars by \cite{orion}.  The objective of this contribution is to provide complete data on all sources from our original sample of 372 protostars as well as the additional targets detected in our images.  Thus, we have included all 410 objects in this contribution.

@@

\section{Results}
\label{sec-results}
\par
In this contribution, we focus on the {\em observed}\ photometric properties of the target protostars, focusing on the fluxes in four bands: the 24~\micron\ from \spitzer/MIPS and the 70, 100, and 160~\micron\ bands from \herschel/PACS instruments..  A series of companion publications, \cite{will}, \cite{furlan}, \cite{erin}, and \cite{pbrpaper} present results and analysis from other scientific investigations and interpretations using these data from the HOPS program.   Our analysis concentrates on the spatial variations in the flux and color distributions across the Orion clouds.  To search for variations in the color and flux distributions, we divide the Orion clouds into 8 distinct sub-regions, shown in Table~\ref{tbl:regions}.  We augment our analysis with previously published results for the PBR sources \cite{pbrpaper}, and sub-mm photometry \citep{thomas}.  We present results from each of the above identified observation quantity below, and collectively discuss the implications for star-formation implied by these observations in Section~\ref{sec:discussion}.

@@

\subsection{Flux ratio distribution}
\label{sec:fratio}
\par
Flux ratios (colors) are better suited for determining differences as a function of source type than the flux itself because flux ratios depend on the shape of the source SED \citep{ali}. We investigated two different flux ratios using the combined Spitzer, HOPS and Gould-Belt photometry: \clra, and \clrb. Figure~\ref{fig:clrclr} shows the flux-ratio vs flux-ratio plots segregated by sub-region for Orion A and Orion B. We do not include data with upper limits for these plots (sources with flag value = 2 in Table~\ref{tbl:obs}). The red symbols show the colors for the sub-region identified in the sub-panel, and the gray symbols show the colors for the entire sample. As with the median of the FD, we note an apparent systematic change in the observed flux ratios relative to the entire population in both Orion~A and Orion~B.
\par
We quantitatively explore this apparent trend further in Figures~\ref{fig:clr1_v_dec}~\&~\ref{fig:clr2_v_dec}. These Figures show the mean and median colors for each of the sub-regions as a function of the mean declination for all sources belonging to the respective sub-region.
\par
In Orion~A, the median and mean colors of the sub-regions become systematically redder towards higher declinations.  In Orion~B, the median and mean colors of the sub-regions becomes systematically redder towards lower declinations.  For both, the vertical bar shows the 1-$\sigma$\ standard deviation of the flux-ratio values as a proxy for the range.  Even with large variation in color, the systematic trends with declination is clearly delineated.

@@

\section{Discussion}
\label{sec:discussion}
\par
The results presented in Section~\ref{sec-results} show the global properties of the HOPS protostar sample.  Since the sub-regions in Orion~A and B are well stratified by declination, differences are easily explored by using declination as a proxy for sub-region.   Indeed, Figures~\ref{fig:mdFlux_v_dec},~\ref{fig:clr1_v_dec},~\&~\ref{fig:clr2_v_dec} show that systematic variations exist in the observed proto-stellar properties as a function of declination in the Orion~A~\&~B clouds. {\bf  We discuss in Sections~\ref{sec:compbias}~\&~\ref{sec:popdiff} two explanations of this observed trend.   Again, we will restrict ourselves to observed quantities only and leave the detailed exploration of inferred properties to companion papers.  We rule out inclination angle (to the observer) as a factor because no reasonable justification can be made for systematic differences in inclination as a function of sub-region in Orion.}

@@

We can now further investigate completeness limit and observational bias via a simple simulation.  The LDN 1641 sub-regions contain the largest number of protostars (see Table~\ref{tbl:regions}) and visual inspection shows the least contamination of the sample from extended nebulosity. Further, fainter protostars are detected in LDN 1641 than in Orion OMC 2/3. Hence, if bright, complex, spatially extended emission from the local dust is responsible for decreased reliability and higher completeness limits for point source photometry, then LDN 1641 provides an ideal contrast to the Orion OMC 2/3 sub-region. The contrast between LDN~1641 and Orion OMC-2/3 is at the extreme ends of the spatial correlation observed in Figures~\ref{fig:mdFlux_v_dec}~\&~\ref{fig:clrclr}.

@@

\subsection{Population differences}
\label{sec:popdiff}
\par
{\bf
We also consider systematic differences in the underlying protostar population as a possible cause for the observed correlations in Figures~\ref{fig:clr1_v_dec}~\&~\ref{fig:clr2_v_dec}.  \cite{will} provide observational evidence for such a scenario.  \cite{will} derived bolometric luminosities and temperatures of our HOPS protostar sample and find statistical differences in the bolometric properties of the protostars as a function of the sub-region in Orion~A~and~B clouds.  The authors argue that these differences track an underlying difference in the star-formation rate across the sub-regions leading to a different mixture of protostars and more evolved objects.  Their findings are supported, in part, by evidence that the fractions for both Class~0 stars and PBRs compared to Class~I~and~II objects are also different in the subregions \citep{pbrpaper}.
\par
\cite{ali} find protostellar envelopes to be the dominant factor affecting the observed colors in the \herschel\ bands.  Given that large, active envelopes are the defining characteristics of Class~0 objects and PBRs, the correlations apparent in Figures~\ref{fig:clr1_v_dec}~\&~\ref{fig:clr2_v_dec} are consistent with the population differences suggested by \cite{will} and \cite{pbrpaper}.  Our data is insufficient to establish causal links for said envelope differences.  However, when taken together, these results support a scenario of systematic differences in the age and/or local environment within the Orion~A~\&~B clouds with the ONC sub-region undergoing the most recent episode of increased star formation or a ramping up of the star formation rate.
\par
Alternatively, changing the size and/or shape of the outflow cavities may also affect the observed \herschel\ colors by allowing more radiation to escape and by the contribution of the warm cavity walls.  We can similarly identify other proto-stellar model parameters investigated by \cite{ali}\ to produce the observed color differences.  We still favor differences in envelopes properties because, as stated earlier, envelopes are the strongest influencers on \herschel\ colors \citep{ali}.
}

\section{Conclusions}
\label{sec:conc}
We explored the mid- to far-infrared properties of the Orion protostars obtained as part of the \herschel Key Program HOPS. Investigations of the observed properties of these protostars reveal apparent correlations as a function of sub-regions in both the Orion A and B clouds.  We investigated the role of bias introduced due to differences in completeness limits in these correlations by artificially introducing a completeness limit in the LDN~1641 sub-region and comparing the resulting changes in median flux ratio to Orion~OMC-2/3 values. These simulations suggest that completeness limit bias alone is not sufficient to explain the observed correlation for \clrb. 
\par
An alternative explanation may be population differences between these sub-regions. {\bf In particular, we find that systematic differences in protostellar envelope densities (correlating with their size and color, see Ali \etal\ 2010) are consistent with population differences suggested by \cite{will} and \cite{pbrpaper}.  In this scenario, we favor the interpretation that \herschel\ colors are influenced by protostellar envelopes over other factors because \cite{ali} find envelopes as the strongest protostellar feature affecting their \herschel\ colors.  Further, this scenario implies that said differences in envelope properties and protostars are likely from differences in the age or the local star-formation environment.  We cannot further differentiate whether evolution or the environment led to such a difference in the protostars.  For example, denser regions may systematically produce denser envelopes which collapse with higher infall rates and may resemble envelopes from a more evolved population with lower density.}
